---
title: "Milestone Project"
author: "Hae Ji Chong"
date: "2024-02-12"
output: pdf_document
---
## Introduction

Money is one thing that is essential in everyone's life.' It can bring joy to someone's life, or it can bring a misery to one's life as well.
I have decided to research on the relationship between income, happiness, and suicidal rates.

## Research questions

In higher income countries, are they more happier than lower income countries?
Does higher income means suicial rate goes down?
If the happiness rate is higher, does that also bring down the suicidal rate?
Has it changed within 3 years period, 2015-2017?
Looking at the data, are we able to predict what the rate would be for the future?

## Approach

I plan to use the datasets to find the relationship among all three. If one goes up, does the other one goes down?
Or do they move in the same direction? 
Also, how has it changed over the time, do they tend to move in same direction every year, are we able to predict the future with these datas?

## How your approach addresses (fully or partially) the problem.

By using this approach, it will state if the income does increase the happiness and brings down the suicidal rate.
And by looking at the data, how strong is the correlation? Are we able to say we see a clear relationship?

## Data  (Minimum of 3 Datasets - but no requirement on number of fields or rows)

https://www.kaggle.com/datasets/unsdsn/world-happiness/data

This shows the happiness level of each country

https://www.kaggle.com/datasets/frankmollard/income-by-country?rvi=1
It has the average income by country
https://www.kaggle.com/datasets/sandragracenelson/suicide-rate-of-countries-per-every-year

It shows the suicide rate of countries per year.
Looking at all three data, I am able to obtain happiness, income, and suicidal rate for 2015- 2018.
I plan on analyzing those four years to see if I can find the relationship.
Income data does have an odd variable at the end, I plan on deleting to make the data cleaner.

## Required Packages
ggplot2
tidyr
dplyr

## Plots and Table Needs

I plan on using scatter plots to see the relationship. I would also do correlation test to get the p-value.
How strong is p-value to state that is positive/negative correlation?

## Questions for future steps

I would need to look into the data sets to see what cleaning needs to be done, such as deleting unnecessary columns.
With the data I have, I need to make sure all three datsets have same countries.

### Part 2
```{r}
library(readxl)
library(dplyr)
```

## How did you import and clean your data?
```{r}
suicide_rate <- read.csv("/Users/my-lo/OneDrive/Desktop/DSC 520/Milestone 1 datasets/suicidal rate by country/suiciderateall.csv")
income_rate <- read_xlsx("/Users/my-lo/OneDrive/Desktop/DSC 520/Milestone 1 datasets/income by country/Income by Country.xlsx", sheet="GNI per capita")
happiness_rate_2015 <- read.csv("/Users/my-lo/OneDrive/Desktop/DSC 520/Milestone 1 datasets/happiness rate by country/2015.csv")
happiness_rate_2016 <- read.csv("/Users/my-lo/OneDrive/Desktop/DSC 520/Milestone 1 datasets/happiness rate by country/2016.csv")
happiness_rate_2017 <- read.csv("/Users/my-lo/OneDrive/Desktop/DSC 520/Milestone 1 datasets/happiness rate by country/2017.csv")
```
I first loaded all the data, for income data I used GNI sheet since I needed look at Gross National Income
```{r}
happiness_rate_2015 <- happiness_rate_2015[order(happiness_rate_2015$Country), ]
happiness_rate_2016 <- happiness_rate_2016[order(happiness_rate_2016$Country), ]
happiness_rate_2017 <- happiness_rate_2017[order(happiness_rate_2017$Country), ]
names(happiness_rate_2015)[names(happiness_rate_2015) == 'Happiness.Score'] <- 'happiness_score_2015'
names(happiness_rate_2016)[names(happiness_rate_2016) == 'Happiness.Score'] <- 'happiness_score_2016'
names(happiness_rate_2017)[names(happiness_rate_2017) == 'Happiness.Score'] <- 'happiness_score_2017'
```
For the happiness data, I decided to put country name by alphabetical order and renamed them to happiness_score_year to make the data more viewer friendly. For income and suicidal rate, country list was already in alphabetical order.
```{r}
happiness_combined<- merge(happiness_rate_2015,happiness_rate_2016,by="Country", all = FALSE)
happiness_combined<- merge(happiness_combined,happiness_rate_2017,by="Country",all = FALSE)
```
I than combined happpiness rate, only included the country that was listd in all three years. I compared 2015 and 2016 first, than compared 2017 with the above file.
```{r}
suicide_rate$Country<-gsub("[^[:alnum:] ]","",suicide_rate$Country)
```
I removed * at the end of the country in suicide rate.
```{r}
new_suicide_rate <- suicide_rate[(as.character(suicide_rate$Country) %in% as.character(income_rate$Country)), ]
new_income_rate <- income_rate[(as.character(income_rate$Country) %in% as.character(new_suicide_rate$Country)), ]
new_happiness_combined <-happiness_combined[(as.character(happiness_combined$Country) %in% as.character(new_income_rate$Country)), ]
new_suicide_rate <- suicide_rate[(as.character(suicide_rate$Country) %in% as.character(new_happiness_combined$Country)), ]
new_income_rate <- income_rate[(as.character(income_rate$Country) %in% as.character(new_suicide_rate$Country)), ]
```
Once again, I only kept the countries that were listed in all the data.
```{r}
happiness_allyear <- new_happiness_combined %>%
  select(Country,happiness_score_2015,happiness_score_2016,happiness_score_2017)
income_allyear <- new_income_rate %>%
  select('Country','2015','2016','2017')
suicide_allyear <-new_suicide_rate %>%
  select(Country,X2015,X2016,X2017)
all_data <-cbind(happiness_allyear,income_allyear,suicide_allyear)
```
With cleaned data that only listed all the countries I needed, I used dplyr and select function to only obtain the columns I need. Each file for happiness, income, and suicide included data for 2015-2017.
```{r}
new_all_df <- subset(all_data, select = -c(5,9))
cols <- c("Country", "happi_2015","happi_2016","happi_2017","Inc_2015","Inc_2016","Inc_2017","Sui_2015","Sui_2016","Suic2017")
colnames(new_all_df) <-cols
```
I now combined them and renamed them to understand what the numbers meant.

## What does the final data set look like?
```{r}
head(new_all_df)
```
Included is my final data set, first six columns.

## What information is not self-evident?

It is not self-evident that all the variables has gone up or down in same direction every year.

I expected the national income has gone up every year, but some countries like Angola has gone down as well.

I cannot assume they have gone up each year, even though most countries have.
Just by lookin at the data, it is not self-evident of the relationship.


## What are different ways you could look at this data?

I plan on using multiple plots to compare them by the year. Scatter plots would be one way.

I plan comparing each country and whether their happiness, income, or suicidal rates change each year.

I could also pick two variables at each time to compare their relationship.

I also want to see if the patterns change in each year. For example, in 2015, was the correlation between happiness and income positive? How about for 2016 and 2017?


## How do you plan to slice and dice the data?

My final data includes all data, I plan on using my sliced data that only includes happiness, income, and suicide rate. In my data, they are named as happiness_allyear, income_allyear, and suicide_allyear

I also plan on slicing the data by year, and plotting all three variables in one graph per year.


## How could you summarize your data to answer key questions?

To answer key questions, I plan on using ANOVA test. I would do anova test among all three variables to see how well it can predict.


## What types of plots and tables will help you to illustrate the findings to your questions?

Scatter plots would help me see the relationship among three variables. With scatter plot, I would draw a line to see how well it fits in a line. 

I also plan on finding residuals using linear regression, and plotting using qqnorm function. Regarding table, summary of ANOVA test would be helpful.

It will tell me how accurately it can predict the relationship.


## Do you plan on incorporating any machine learning techniques to answer your research questions? Explain.

I plan on incorporating linear regression to see if I can draw a line through the relationships. 

I also plan on using summary of the linear regression to see how strong the correlation is.

Is it strong enough to say it is related? I would also look at P-value to see if it is less than 0.05 to see their relationship.


## What questions do you have now, that will lead to further analysis or additional steps?

For further steps, I need to do further analysis to see which graphs or plots make the most sense to accurately show three relationships among variables. 

I plan on having multiple plots in one graph to see the relationship, is that powerful and does that accurately measure the relationship?

And, does that accurately align with what my linear regression method?

### Part 3

## Introduction.
Money is a big part of someone's life. It can make someone happy, or someone sad.
If people live in a country with higher income, and if that can lead to happiness, does it make sense to move to that area?

## The problem statement you addressed.
Problem statement I addressed was what are the relationship among all three variables.
For those countries that are wealthier than others, are they more happier than those countries thare were poor?
And because they have money, does the depression rate also goes down? 
If the happiness goes up, do suicidal rate also goes down in same rate?
I want to see if there is a trend among those three variables.

## How you addressed this problem statement
```{r}
library(tidyverse)
ggplot(data=new_all_df, aes(x=Inc_2015,y=Sui_2015,))+geom_point()+xlab("Income 2015")+ylab("Suicide rate 2015")+ggtitle("Income vs suicide rate for 2015")
ggplot(data=new_all_df, aes(x=Inc_2016,y=Sui_2016,))+geom_point()+xlab("Income 2016")+ylab("Suicide rate 2016")+ggtitle("Income vs suicide rate for 2016")
ggplot(data=new_all_df, aes(x=Inc_2017,y=Suic2017,))+geom_point()+xlab("Income 2017")+ylab("Suicide rate 2017")+ggtitle("Income vs suicide rate for 2017")

ggplot(data=new_all_df, aes(x=Inc_2015,y=happi_2015,))+geom_point()+xlab("Income 2015")+ylab("Happiness rate 2015")+ggtitle("Income vs Happiness rate for 2015")
ggplot(data=new_all_df, aes(x=Inc_2016,y=happi_2016,))+geom_point()+xlab("Income 2016")+ylab("Happiness rate 2016")+ggtitle("Income vs Happiness rate for 2016")
ggplot(data=new_all_df, aes(x=Inc_2017,y=happi_2017,))+geom_point()+xlab("Income 2017")+ylab("Happiness rate 2017")+ggtitle("Income vs Happiness rate for 2017")
```
I addresssed this problem by analyzing all three variables. I have first drawn scatterplots of Income vs. happiness and income vs. suicidal rate.

```{r}
new_all_df$Inc_2015<-as.numeric(new_all_df$Inc_2015)
new_all_df$Inc_2016<-as.numeric(new_all_df$Inc_2016)
new_all_df$Inc_2017<-as.numeric(new_all_df$Inc_2017)
```
Because the income data type was "character", I had to change them to numeric for further analysis
```{r}
income_lm_2015 <- lm(formula=happi_2015~Inc_2015+Sui_2015, data=new_all_df)
summary(income_lm_2015)
income_lm_20152 <- lm(formula=happi_2015~Inc_2015, data=new_all_df)
summary(income_lm_20152)
income_lm_2016 <- lm(formula=happi_2016~Inc_2016+Sui_2016, data=new_all_df)
summary(income_lm_2016)
income_lm_2017 <- lm(formula=happi_2017~Inc_2017+Suic2017, data=new_all_df)
summary(income_lm_2017)
suicidal_lm <-lm(formula = Sui_2015~Inc_2015+happi_2015, data=new_all_df)
summary(suicidal_lm)
```
Once I drew the scatterplot, I also did a linear regressison to see if I see the relationship of happines with income and suicidal rate. 
When I did all 3 years, I noticed they were all similar.
Because they were similar, I decided to focus the anlaysis of one year, which I picked as 2015.
I did the linear regression of happiness just with income factor, without suicidal rate as well.
I also did a linear regression of suicidal rate with income and happiness factor for the year of 2015.
```{r}
res <- resid(income_lm_2015)
plot(fitted(income_lm_2015),res)
abline(0,0)
qqnorm(res)
qqline(res)
```

I did plot the residual to see if I can say they are normal or not.
With how well the line fits into each points, I can say it is normal.

```{r}
cor.test(new_all_df$happi_2015,new_all_df$Inc_2015)
cor.test(new_all_df$Sui_2015,new_all_df$Inc_2015)
cor.test(new_all_df$Sui_2015,new_all_df$happi_2015)
```
I also did a correlation test to see the correlation.

## Analysis.
By doing the analysis, I have to say there is a positive relationship between income and happiness because of high correlation.
However, there was no strong relationship between income and suicidal rate, and suicidal rate and happiness.
When I look at the year of 2015, linear regression of R squred came back as 52.07%. 
With income and suicidal rate, we can predict on happiness 52.07% of the time.
With linear regression of happiness just on the income factor, R squared was 51.53%. 
Therefore, adding the suicidal rate with income factor increased the prediction rate of happiness.
Linear regression of suicidal rate was not significant due to p-value higher than 0.05. 
With R squared being 0.063, we can only predict 6% of the time by knowing happiness and income.

## Implications. 
This could imply that people tend to be happier in countries that are wealthier since they have higher income.
Moving to a wealthier country and having an opportunity to have a high income could lead to happiness of one's life.
However, high happiness rate does not imply lower suicidal rate.
Some countries could have the highest happiness ranking, but they could also have high suicidal rate.
With my data, we can predict happiness over 52% of the time by knowing income and suicidal rate, but it is hard to predict suicidal rate just by knowing income and happiness rate.

## Limitations.
My data is divided among countries. In countries like the U.S., some states are wealthier than the other states.
For someone to learn more about the relationship, they could analyze within each countries.
One other limitation is how they have collected the data of happiness. 
Income and suicidal rate could be obtained objectively, but happiness is subjective. collecting the data on holiday season could be different than non holiday season as well.
If we were able to find an objective way to obtain happiness, the data could be more reliable and accurate.

## Concluding Remarks
With my data, I can say there are some relationship betweeen happines and income. 
It's hard to assume that happiness is caused by high income since there could be other factors between the two variables.
Also, since the relationship between suicidal and income was not high enough, we can't say people with low income gets more depressed than people with high income.
To build on my analysis, it would be interesting to see what other factors can have impact on one's happines.
Instead of just looking at income as in GNI, gross national income, it would be interesting to include GDP, gross domestic product, in part of an analysis as well.
